---
title: Mapping Trips to Aunt Zerviah's
execute:
  echo: false
---

*"They have all with the exception of Mother Lib and me gone to Aunt Zerviah to spend the evening I tired out every day by dark as I have no place to lie down and rest during the day. Lib and I am going to retire early"*

**Celestia Webster's Diary - November 13, 1867**

Map of Trips to Aunt Z just points

```{r}
"library(tidyverse)
library(sf)
library(leaflet)
library(paletteer)

zlocations <- read_csv(z_locations)

zlocations_sf <- st_as_sf(zlocations,
                         coords = c(lng, lat),
                         crs = 4326)
pal <- colorFactor(
  palette = as.character(paletteer_d(colorBlindness::paletteMartin)),
  domain = zlocations_sf$location)
                 leaflet() |> 

leaflet() |> 
  addTiles() |>
  addCircleMarkers(data = zlocations_sf,
                   fillOpacity = 0.5,
                   radius = 3) |>
  addLegend(bottomleft, pal = pal, values = zlocations_sf$location)"
```

## Map of Trips to Aunt Zerviah - One line for each individual

```{r}
library(tidyverse)
library(sf)
library(leaflet)
library(paletteer)

## read data

zmovements <-read_csv("z_movements.csv")
zlocations <-read_csv("z_locations.csv")

#sf object

zlocations_sf <- st_as_sf(zlocations,
                         coords = c("lng", "lat"),
                         crs = 4326)
#joining geocoded data to attribute data and then turning it into an sf object

zmovements_geo <- left_join(zmovements, zlocations,
                           by = join_by(destination == location))
zmovements_sf <- st_as_sf(zmovements_geo,
                         coords = c("lng", "lat"),
                         crs = 4326)
#using the group by and summarize workflow 
zmovements_lines <- zmovements_sf |> 
  arrange(date) |>
  group_by(person) |>
  summarise(do_union = FALSE) |>
  st_cast("LINESTRING")

#leaflet map
pal <- colorFactor(
  palette = as.character(paletteer_d("colorBlindness::paletteMartin")),
    domain = zmovements_lines$person)

leaflet() |> 
  addTiles() |>
  addPolylines(data = zmovements_lines,
               color = ~pal(person),
               opacity = 0.8,
               label = ~person) |> 
  addCircleMarkers(data = zlocations_sf,
                   fillOpacity = 0.5,
                   radius = 5,
                   label = ~location) |> 
  addLegend("bottomleft", pal = pal, values = zmovements_lines$person)


```

## Map of Trips to Aunt Zerviah - One line for each individual

```{r}
"library(sf)
library(tidygeocoder)
library(USAboundaries)
library(tidyverse)
library(leaflet)
library(paletteer)

zlocations <- read.csv(z_locations.csv)
zmovements <- read.csv(z_movements.csv)

zlocations_sf <- st_as_sf(zlocations,
                         coords = c(lng, lat),
                         crs = 4326)
zmovements_wide <- zmovements |> 
  arrange(date) |> 
  group_by(person) |>
  ungroup() |>
  filter(!is.na(start)) |>
    rowid_to_column(journeyid)

zmovements_long <- zmovements_wide |> 
  pivot_longer(
    cols = start:destination,
    names_to = type,
    values_to = location
  )

zmovements_long_sf <- zmovements_long |> 
  left_join(zlocations, by = location) |> 
  st_as_sf(coords = c(lng, lat), crs = 4326) |> 
  group_by(journey_id) |> 
  summarise(do_union = FALSE) |> 
  st_cast(LINESTRING)

journeys_sf <- left_join(zmovements_long_sf, zmovements_wide, by = journey_id)
journeys_sf

journeys_sf <- journeys_sf |> 
  mutate(label = paste (
    person, <br/>,
    start, to, destination, <br>,
    year(date),day(date), month(date, label = TRUE)))

pal <- colorFactor(
  palette = as.character(paletteer_d(colorBlindness::paletteMartin)),
  domain = journeys_sf$person)

leaflet() |> 
  addTiles() |>
  addPolylines(data = journeys_sf,
               color = ~pal(person),
               opacity = 0.8,
               label = ~person) |> 
  addCircleMarkers(data = zlocations_sf,
                   fillOpacity = 0.5,
                   radius = 5,
                   label = ~location) |> 
  addLegend(bottomleft, pal = pal, values = journeys_sf$person)"
```
